DATA 621 - Business Analytics and Data Mining

### 2020-12-01 Final_project/Final Project Paper - Group 2.Rmd
Add first run at rules (country) based linear model

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```
```{r loadLibraries}
library(knitr)
library(dplyr)
library(dtplyr)
library(stringr)
library(ggplot2)
library(scales)
library(data.table)
```
```{r loadData}
dataHeader <- "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/"
ca_path <- paste0(dataHeader, "CAvideos.csv")
us_path <- paste0(dataHeader, "USvideos.csv")
gb_path <- paste0(dataHeader, "GBvideos.csv")
fr_path <- paste0(dataHeader, "FRvideos.csv")
de_path <- paste0(dataHeader, "DEvideos.csv")
droppedFields <- c('description', 'thumbnail_link')
ca_set <- fread(ca_path, drop = droppedFields)
us_set <- fread(us_path, drop = droppedFields)
gb_set <- fread(gb_path, drop = droppedFields)
fr_set <- fread(fr_path, drop = droppedFields)
de_set <- fread(de_path, drop = droppedFields)
nobsUS <- nrow(us_set)
nobsCA <- nrow(ca_set)
nobsGB <- nrow(gb_set)
nobsFR <- nrow(fr_set)
nobsDE <- nrow(de_set)
nobsTot <- nobsUS + nobsCA + nobsGB + nobsFR + nobsDE
```

@@

Given the models, we will compare the RMSE and MAE on a special holdout set and
the AIC on the training set and will select the model that performs best as the
winner for this paper. We do not expect it to outperform more sophisticated
models.

@@

## Feature Engineering

## Model Building

@@

```{r loadLibraries}
```
```{r loadData}
```
```{r ca_data_prep}
```
```{r de_data_prep}
```
```{r fr_data_prep}
```
```{r gb_data_prep}
```
```{r us_data_prep}
```
```{r merge_datasets}
```
```{r mungeData}
```
```{r summaryStats}
```
```{r factorPlot1}
```
```{r factorPlot2}
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```
```{r loadLibraries}
library(doParallel)
library(knitr)
library(dplyr)
library(dtplyr)
library(stringr)
library(ggplot2)
library(scales)
library(caret)
library(data.table)
```
```{r loadData}
dataHeader <- "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/"
ca_path <- paste0(dataHeader, "CAvideos.csv")
us_path <- paste0(dataHeader, "USvideos.csv")
gb_path <- paste0(dataHeader, "GBvideos.csv")
fr_path <- paste0(dataHeader, "FRvideos.csv")
de_path <- paste0(dataHeader, "DEvideos.csv")
droppedFields <- c('description', 'thumbnail_link')
ca_set <- fread(ca_path, drop = droppedFields)
us_set <- fread(us_path, drop = droppedFields)
gb_set <- fread(gb_path, drop = droppedFields)
fr_set <- fread(fr_path, drop = droppedFields)
de_set <- fread(de_path, drop = droppedFields)
nobsUS <- nrow(us_set)
nobsCA <- nrow(ca_set)
nobsGB <- nrow(gb_set)
nobsFR <- nrow(fr_set)
nobsDE <- nrow(de_set)
nobsTot <- nobsUS + nobsCA + nobsGB + nobsFR + nobsDE
```

@@

Given the models, we will compare the RMSE and MAE on a special holdout set and
will select the model that performs best as the winner for this paper. We do not
expect it to outperform more sophisticated models.

@@

## Feature Engineering & Selection
Since we are not factoring in time, it is incorrect to use all the observations.
Therefore, we will extract the latest observation by video by country and use
this subset.

```{r subsetExtraction}
# Select the row numbers of the first entry of the latest trending date by title
# and by country. There are 34 duplicates. Use this as our restricted data set
# corresponding to the most recent view count. The inner set of brackets gets
# the row number (called V1) and the outer set is a simple subset by those row
# numbers.
fullExtract <- full_set[
  full_set[, .I[which.max(trending_date)], by = c('title', 'country')]$V1
  ]
```

A feature that we will also consider is `lkratio`: the ratio of likes to
dislikes. This will help ascertain if it is the magnitude or the preponderance
which has a greater effect. As there may be some rare videos with no dislikes,
the ratios will be set to 1. To capture the effect of likes with no dislikes, an
indicator variable will be created or those "perfect" videos.

```{r addFeatures}
fullExtract[, `:=`(perfect =ifelse(likes > 0 & dislikes == 0, TRUE, FALSE),
                   lkratio = ifelse(dislikes == 0, 1, likes / dislikes))]
```

## Model Building
We will first separate 20% of the data as a true holdout set. It is on this data
that our models will be compared.

```{r holdOut}
set.seed(617)
seenIDX <- createDataPartition(fullExtract$views, p = 0.8)$Resample1
seenSet <- fullExtract[seenIDX, ]
hideSet <- fullExtract[-seenIDX, ]
```

We will train models on the remaining 80% of the data.

### Model 1: Linear Regression
Keeping country as a predictor is of less interest. It will most likely track
some measure of population. It is of more interest to see if there are
differences in the factors between countries. To capture this by interactions
would be combinatorially unfeasible. Instead, five models will be run, each on
a country specific subset. This is akin to a manual implementation of a rules
based regression model

```{r modelLinear1, cache = 2L}
# Using stepAIC means no cross-validation. Train on entire dataset.
trC <- trainControl(method = 'none', allowParallel = TRUE)
cl <- makePSOCKcluster(5)
registerDoParallel(cl)
set.seed(181)
for (ct in unique(fullExtract$country)) {
  linMod <- train(views ~ likes + dislikes + comments + tags + category_id +
                    comments_disabled + ratings_disabled +
                    video_error_or_removed + perfect + lkratio,
                  data = fullExtract[country == ct],
                  family = gaussian(link = 'identity'), direction = 'both',
                  method = 'glmStepAIC', trace = 0, trControl = trC)
  assign(paste0('linMod', ct), linMod)
}
stopCluster(cl)
lM1US <- linModUS$finalModel
lM1CA <- linModCA$finalModel
lM1GB <- linModGB$finalModel
lM1FR <- linModFR$finalModel
lM1DE <- linModDE$finalModel
clM1US <- linModUS$finalModel$coefficients
clM1CA <- linModCA$finalModel$coefficients
clM1GB <- linModGB$finalModel$coefficients
clM1FR <- linModFR$finalModel$coefficients
clM1DE <- linModDE$finalModel$coefficients
nclM1US <- names(clM1US)
nclM1CA <- names(clM1CA)
nclM1GB <- names(clM1GB)
nclM1FR <- names(clM1FR)
nclM1DE <- names(clM1DE)
nnclM1US <- length(nclM1US)
nnclM1CA <- length(nclM1CA)
nnclM1GB <- length(nclM1GB)
nnclM1FR <- length(nclM1FR)
nnclM1DE <- length(nclM1DE)
length(nclM1US) <- 22
length(nclM1GB) <- 22
length(nclM1FR) <- 22
length(nclM1DE) <- 22
```

\footnotesize
```{r m1Table}
kable(data.frame(cbind(nclM1CA, nclM1DE, nclM1US, nclM1FR, nclM1GB)),
      caption = "Model Coefficients by Country")
```
\normalsize

The most parsimonious model is the one for Great Britain with only `r nnclM1GB`
parameters. The most complex is the one for Canada with `r nnclM1CA`.

@@

```{r loadLibraries}
```
```{r loadData}
```
<!--  Data Exploration -->
```{r ca_data_prep}
```
```{r de_data_prep}
```
```{r fr_data_prep}
```
```{r gb_data_prep}
```
```{r us_data_prep}
```
```{r merge_datasets}
```
```{r mungeData}
```
```{r summaryStats}
```
```{r factorPlot1}
```
```{r factorPlot2}
```

<!-- Feature Engineering & Selection -->
```{r subsetExtraction}
```
```{r addFeatures}
```

<!-- Model Building -->
```{r holdOut}
```
```{r modelLinear1}
```
```{r m1Table}
```
